---
title: "Building energy prediction"
author: "Athanasios Pliousis"
date: "21/6/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
library(tidyverse)
library(lubridate)
library(scales)
library(ggrepel)
library(patchwork)
library(broom)
```


## Reading the data

```{r Import data}
df_train <- read_csv("train.csv")
df_test <- read_csv("test.csv")
building_meta <- read_csv("building_metadata.csv")
weather_test <- read_csv("weather_test.csv")
weather_train <- read_csv("weather_train.csv")
# sample_submission <- read_csv("sample_submission.csv")
```


```{r Unique identification}
df_train$building_id %>% unique() %>% NROW() # 1449 buildings
building_meta$site_id %>% unique() # 15 unique sites
building_meta$primary_use %>% unique() # 16 unique uses
```


```{r Making a square_meters variable}
building_meta <- building_meta %>% 
    mutate(square_meter = square_feet * 0.09290304) 
# 1 square foot = 0.09290304 square meters
```

```{r}
joined <- df_train %>% full_join(building_meta, by = 'building_id')
rm(df_train)
```



```{r}
# Correction to units for site 0 electric meter
# posted in ashrae-energy-prediction 7 months ago [gold]
# 
# Due to a miscommunication, the electric meter readings for site 0 were not properly converted to units of kWh and are in kBTU. We apologize for the error and the resulting inconvenience.
# 
# The units of the electric meter readings for site 0 will not be updated (i.e., they will remain as is), since this could potentially disrupt the workflows and models of many teams. For models that are sensitive to units/absolute values, the issue can be avoided with the right conversion factors. Multiply by 0.2931 to get to model inputs into kWh like the other sites, and 3.4118 to get back to kBTU for scoring.
joined <- joined %>% 
    mutate(meter_reading = if_else(site_id == 0 & meter == 0, 
                                   meter_reading *  0.2931,
                                   meter_reading))
```


How many building have each meter?
```{r}
electric_buildings <- joined %>% filter(meter == 0) %>% # 0:electricity
    count(building_id) %>% 
    pull(building_id)   # 1413 buildings all 15 sites
chill_buildings <- joined %>% filter(meter == 1) %>% # 1:chilledwater
    count(building_id) %>% 
    pull(building_id)   # 498 buildings
steam_buildings <- joined %>% filter(meter == 2) %>% # 2:steam
    count(building_id) %>% 
    pull(building_id)   # 324 buildings
hotwater_buildings <- joined %>% filter(meter == 3) %>% # 3:hotwater
    count(building_id) %>% 
    pull(building_id)   # 145 buildings
```



```{r}
joined$timestamp <- joined$timestamp %>% lubridate::ymd_hms()
```


Create the variable energy per square meter to be able to compare energy use on a common characteristic among buildings.

```{r}
joined <- joined %>% mutate(energy_sqm = meter_reading / square_meter)
```

```{r}
joined$energy_sqm %>% summary()
```

The max energy per sqm is extremely high for the max value. The until the 3rd quantile it is below 0.0311.
Lets look at what is happening after that.

```{r}
joined %>% filter(energy_sqm == max(energy_sqm))
```


```{r}
joined$energy_sqm %>% quantile(c(0.95, 0.999))
```
Let first find buildings that are over a threshold of energy_sqm > 5 rather the 99.9% percentile value. And we will see if we 
change it after.

```{r}
building_outliers <- joined %>% filter(energy_sqm > 5) %>% distinct(building_id) %>% arrange(building_id)
```


```{r}
outlier_building_summary <- joined %>% 
  filter(building_id %in% pull(building_outliers)) %>%
  group_by(building_id) %>% 
  summarise(summaries = list(tidy(summary(energy_sqm)))) %>% 
  unnest(summaries) %>% 
  arrange(median)
  
outlier_building_summary
```

Buildings 1099 and 778 are way off in terms of energy per square meter. I will examine them in isolation but lets
first do the histograms of all the energy_sqm values broken down in a step by step fashion. The breaks are arbitrary
put by me looking at the data and making sense of the spread of the values.


```{r}
energ_hist_0_025 <- joined %>% filter(energy_sqm > 0, energy_sqm < 0.25) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 0.001)
energ_hist_025_1 <- joined %>% filter(energy_sqm > 0.25, energy_sqm < 1) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 0.001)
energ_hist_1_5 <- joined %>% filter(energy_sqm > 1, energy_sqm < 5) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 0.1)
energ_hist_5_20 <- joined %>% filter(energy_sqm > 5, energy_sqm < 20) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 0.1)
energ_hist_20_30 <- joined %>% filter(energy_sqm > 20, energy_sqm < 30) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 0.1)
energ_hist_30_80 <- joined %>% filter(energy_sqm > 30, energy_sqm < 80) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 1)
energ_hist_80_over <- joined %>% filter(energy_sqm > 80) %>% ggplot(aes(energy_sqm)) + geom_histogram(binwidth = 1)
(energ_hist_0_025 | energ_hist_025_1 | energ_hist_1_5 ) / (energ_hist_5_20 | energ_hist_20_30 | energ_hist_30_80) / energ_hist_80_over
```



```{r }
joined %>% filter(building_id == 778) %>% 
  ggplot(aes(timestamp, energy_sqm, color = as.factor(meter))) + 
  geom_point(alpha = 0.9) + 
  theme(legend.position = "top") +
  scale_x_datetime(labels = date_format("%b"), date_breaks = "1 month") +  
  #theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~meter, scales = "free_y")
```
It is obvious that around October there is something going one that changes completely the behavior of usage in meter 1: chilledwater. The difference is extreme. If we had data over one year i could check to see if it is a event related pattern. I should check the test data to see if it is repeating.

Lets first check the electricity consumption. which seems to have relatively stable baseline consumption and hourly peaks. 
The low values which i consider as the night consumption (and i defined previously as the base line) seem elevated
between January and mid March and again at the end of October and through January.

In the modeling phase a smoothing over every hour in the week or in a month multiple could summarize the repeating trend. 


```{r }
joined %>% filter(building_id == 778, 
                  meter == 0) %>% 
                  #timestamp >= min(timestamp) + lubridate::ddays(2), timestamp <= min(timestamp) + lubridate::ddays(4)) %>%  
  ggplot(aes(timestamp, energy_sqm)) + 
  geom_line(alpha = 0.5)  +
  scale_x_datetime(labels = date_format("%d-%m--%H"), date_breaks = "1 month") +  
  theme(axis.text.x = element_text(angle = 90))
```
Lets check what that outlier is. 

```{r}
joined %>% filter(building_id == 778, meter == 0, timestamp >= "2016-08-24", timestamp < "2016-08-25") %>% 
  ggplot(aes(timestamp, energy_sqm)) + 
  geom_line(alpha = 0.5)
```
It is probably a power outage? I will probably put the same consumption of the previous day in that time duration to fill the gap.

```{r}
joined %>% filter(building_id == 778, meter == 0, timestamp >= "2016-08-23", timestamp < "2016-08-25") %>% 
  select(timestamp, meter_reading, energy_sqm)
```
			
The two problematic values are: 
timestamp   meter_readinge  energy_sqm

2016-08-24 14:00:00	0.000	0.00000000		
2016-08-24 15:00:00	0.000	0.00000000	

So i will replace them with
2016-08-23 14:00:00	327.783	0.03256654		
2016-08-23 15:00:00	327.033	0.03249203


```{r}
# The row with the problem
d24_08_1400 <- joined %>% filter(building_id == 778, meter == 0, timestamp == ymd_hms("2016-08-24 14:00:00"))
d24_08_1500 <- joined %>% filter(building_id == 778, meter == 0, timestamp == ymd_hms("2016-08-24 15:00:00"))
mr23_08_1400 <-  joined %>% 
  filter(building_id == 778, meter == 0, timestamp == ymd_hms("2016-08-23 14:00:00")) %>% pull(meter_reading)
mr23_08_1500 <-  joined %>% 
  filter(building_id == 778, meter == 0, timestamp == ymd_hms("2016-08-23 15:00:00")) %>% pull(meter_reading)
en_sqm23_08_1400 <-  joined %>% 
  filter(building_id == 778, meter == 0, timestamp == ymd_hms("2016-08-23 14:00:00")) %>% pull(energy_sqm)
en_sqm23_08_1500 <-  joined %>% 
  filter(building_id == 778, meter == 0, timestamp == ymd_hms("2016-08-23 15:00:00")) %>% pull(energy_sqm)
joined <- joined %>%
  mutate(meter_reading = replace(meter_reading,
             # having already found  where the problematic value is a am setting it          
                                 meter_reading == d24_08_1400 %>% pull(meter_reading),
          # and replacing the meter_reading value with the same value of the previous day
                                 mr23_08_1400),
         
         meter_reading = replace(meter_reading,
                                 meter_reading == d24_08_1500 %>% pull(meter_reading),
                                 mr23_08_1500),
         
         # Same thing and timestamp for replacing the energy_sqm value. 
         energy_sqm = replace(energy_sqm,
                              energy_sqm == d24_08_1400 %>% pull(energy_sqm), 
                              en_sqm23_08_1400),
         
         energy_sqm = replace(energy_sqm,
                              energy_sqm == d24_08_1500 %>% pull(energy_sqm), 
                              en_sqm23_08_1500)
         )
  
```






```{r }
joined %>% filter(building_id == 1099) %>% 
  ggplot(aes(timestamp, energy_sqm, color = as.factor(meter))) + 
  geom_point(aes(alpha = 0.9)) + 
  theme(legend.position = "none")
```


```{r }
joined %>% 
  filter(energy_sqm >= quantile(energy_sqm, 0.95), !(building_id %in% c(778, 1099))) %>% 
  ggplot(aes(timestamp, energy_sqm, color = as.factor(building_id))) + 
  geom_point(alpha = 0.5) + 
  theme(legend.position = "none")
```